Search-based Synthesis of Network Intrusion Detection Rules

### 2019-09-23 paper/paper.tex
pass on section II

This section 1) shows an example rule of Suricata and briefly explains
the format of Suricata rules, 2) presents an example of
a network attack, and 3) illustrates how \tname{} performs to
synthesize the rule that would capture the attack.

@@

\Mar{Lucas/Guilherme, verifique se isto esta correto}
Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
the Open Information Security Foundation (OISF)~\cite{oisf}. A
Suricata rule is divided in three parts---action, header, and rule
options~\cite{suri-rule-format}. The action part appears as the first
word in the rule description. In this case, \CodeIn{alert}. An action
denotes the task that needs to be executed if the rule pattern is
satisfied. In this case, a message will be sent to sys admins if the
rule pattern is satisfied. Other options of actions include pass,
drop, and reject. The header comes after the action in the rule
description. It restricts the information flow covered by the
rule. For this rule, the header is \CodeIn{tcp \$HOME\_NET any ->
  \$EXTERNAL\_NET any}, which instructs Suricata to inspect
\CodeIn{tcp} traffic flowing from any port in the home network to any
other address outside the home network. The variables
\CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are
configurable. The rule options come after the header. It is a
semi-colon-separated sequence of properties\Mar{should we call it an
  option or property?}, where each property is a key-value pair
describing a distinct characteristic of the attack.

@@

The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
works as follows. First, a client sends a SYN packet to the server,
requesting a connection. Second, the server responds with a SYN-ACK
packet to the client. Third, the client responds with an ACK message
and the connection is established. Aware of the protocol, an attacker
sends multiple SYN packets to different ports of a server, often using
fake IP addresses. Without proper protection, the server accepts the
connection requests and eventually legitimate connection requests
cannot be satisfied due to resource exhaustion.

Figure~\ref{fig:synflood-example} shows a Suricata rule to capture
this attack. The critical parts are the option \CodeIn{flags: S,12},
which identifies a SYN packet and the option \CodeIn{threshold: type
  both, track by\_dst, count 5000, seconds 5}, indicating that a high
volume of such packets should be requested in a short period of time.

@@

\Mar{Guilherme/Lucas (CRI): we need to show how \tname{} works on this
  example. What is the input? How long does it take to generate the
  rule? Is the rule generate exactly the same?}

>>>>

This section 1)~shows an example rule of Suricata, 2)~briefly explains
the format of Suricata rules, 3) presents an example of a network
attack, and 4) illustrates how \tname{} performs to synthesize the
rule that would capture the attack.

@@

\Mar{Lucas/Guilherme, verifique se isto esta correto}
Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
the Open Information Security Foundation (OISF)~\cite{oisf}. A
Suricata rule is divided in three parts---action, header, and rule
options~\cite{suri-rule-format}. The action part appears as the first
word in the rule description. In this case, \CodeIn{alert}. An action
denotes the task that needs to be executed if the rule pattern is
satisfied. In this case, a message will be sent to sys admins if the
rule pattern is satisfied. Other actions include pass, drop, and
reject. The header comes after the action in the rule description. It
restricts the information flow covered by the rule. For this rule, the
header is \CodeIn{tcp \$HOME\_NET any -> \$EXTERNAL\_NET any}, which
instructs Suricata to inspect \CodeIn{tcp} traffic flowing from any
port in the home network to any other address outside the home
network. The variables \CodeIn{\$HOME\_NET} and
\CodeIn{\$EXTERNAL\_NET} are configurable. The rule options come after
the header. It is a semi-colon-separated sequence of options, where
each property is a key-value pair describing a distinct characteristic
of the attack.

@@

The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
works as follows. First, a client sends a SYN packet to the server,
requesting a connection. Second, the server responds with a SYN-ACK
packet to the client. Third, the client responds with an ACK message
and the connection is established. Aware of the protocol, an attacker
sends multiple SYN packets to different ports of a server, often using
fake IP addresses.\Mar{Are you saying that there is no timer that the
  server starts when it sends a SYN-ACK message?} Without proper protection, the server accepts the
connection requests and eventually legitimate connection requests
cannot be satisfied due to resource exhaustion.

Figure~\ref{fig:synflood-example} shows a Suricata rule to capture
this attack. The critical parts are the option \CodeIn{flags: S,12},
which identifies a SYN packet and the option \CodeIn{threshold: type
  both, track by\_dst, count 5000, seconds 5}, indicating that a high
volume of such packets should be requested in a short period of
time.\Mar{please, explain the other fields}

@@

\Mar{Guilherme/Lucas (CRI): we need to show how \tname{} works on this
  example. What is the input? How long does it take to generate the
  rule? Is the rule generated exactly the same? Please explain...}
